---
title: 'Data-Cleaning'
author: "Joshua Lin"
format: 
  html: 
    embed-resources: true
    toc: true
    code-fold: true
---

```{r}
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(lubridate))

airbnb <- read.csv("Data/raw-data/nyc-airbnb-listings-2024.csv")

# First 5 rows in a dataset
# head(airbnb)

airbnb <- airbnb |> select(-license, -last_review)
colnames(airbnb)
```

> Note: There are 18 columns, after dropping columns "license" and "last_review", where they can be divided into 4 major categories:
> 
> 1. Listing Details
> 
> 2. Host Information
> 
> 3. Location
> 
> 4. Availability & Reviews

Upon glancing at the dataset, I can notice that there are quite a substantial of missing values. 

For example regarding missing values in columns "name" and "host_name", I decided to fill the missing values with "Unknown". Additionally as a result of "price" being essential for analysis, I removed listings where the price is missing or 0. 

```{r}
# Handling missing data
airbnb$name[is.na(airbnb$name)] <- "Unknown"
airbnb$name[is.na(airbnb$host_name)] <- "Unknown"
airbnb$reviews_per_month[is.na(airbnb$reviews_per_month)] <- 0

airbnb <- airbnb |> filter(!is.na(price) & price > 0)

# Removing duplicates from dataset
airbnb <- airbnb |> distinct()

# Making sure data types are consistent
airbnb$price <- as.numeric(airbnb$price)
airbnb$minimum_nights <- as.numeric(airbnb$minimum_nights)
airbnb$review_per_month <- as.numeric(airbnb$reviews_per_month)

# Using IQR to handle extreme outliers in the "price" col
Q1 <- quantile(airbnb$price, 0.25, na.rm = TRUE)
Q3 <- quantile(airbnb$price, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
price_upper_limit <- min(Q3 + 1.5 * IQR, 1000)
airbnb <- airbnb %>% filter(price <= price_upper_limit)

airbnb <- airbnb %>% filter(minimum_nights <= 365)

airbnb <- airbnb %>% filter(reviews_per_month <= 10)

write.csv(airbnb, "Data/processed-data/nyc-airbnb-listings-2024-cleaned.csv")

# any(is.na(airbnb))
```
> We can conclude that our new dataset has no missing values. Therefore, it is now cleaned and ready for analysis. 