---
title: 'Exploratory Data Analysis'
author: "Joshua Lin"
format: 
  html: 
    embed-resources: true
    toc: true
    code-fold: true
---

```{r}
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
suppressMessages(library(ggplot2))

airbnb <- read.csv("Data/processed-data/nyc-airbnb-listings-2024-cleaned.csv")

# head(airbnb)

# Gives the summary statistics
summary(airbnb)
```

```{r}
# Price Distribution
ggplot(airbnb, aes(x = price)) +
    geom_histogram(bins = 50, fill = "steelblue", color = "white") +
    scale_x_continuous() +
    labs(
        title = "Distribution of Airbnb Prices in NYC",
        x = "Price (USD per night)",
        y = "Frequency"
    ) +
    theme_minimal()
```

This histogram depicts how Airbnb listing prices are distributed across NYC.

```{r}
# Distribution of Room Types

# Count room types and reorder
room_type_counts <- airbnb %>%
    count(room_type)

# Create the bar chart
ggplot(room_type_counts, aes(x = reorder(room_type, n), y = n, fill = room_type)) +
    geom_bar(stat = "identity", color = "black") +
    geom_text(aes(label = n), vjust = -0.5, size = 4) +
    scale_fill_manual(values = c("#2E8B57", "#1C4E80", "#6B6B6B", "#FFD700", "#A73535")) +
    labs(
        title = "Distribution of Room Types in NYC Airbnb Listings",
        x = "Room Type",
        y = "Number of Listings",
        fill = "Room Type"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none"
    )
```

This bar chart visualizzes the number of Airbnb listings for each room type in NYC. Not that "entire home/apt" is the most common listing type. 

```{r}
# Count listings by borough
borough_counts <- airbnb %>%
    count(neighbourhood_group, sort = TRUE)

# Calculate percentages for labeling
borough_counts <- borough_counts %>%
    mutate(percentage = round((n / sum(n)) * 100, 1))

# Custom color palette
nyc_palette <- c(
    "Bronx" = "#A73535", "Brooklyn" = "#B5A21F",
    "Manhattan" = "#2E8B57", "Queens" = "#1C4E80", "Staten Island" = "#A86EAA"
)

# Plot borough distribution
ggplot(borough_counts, aes(x = reorder(neighbourhood_group, n), y = n, fill = neighbourhood_group)) +
    geom_bar(stat = "identity", width = 0.6, color = "black") +
    scale_fill_manual(values = nyc_palette) +
    labs(
        title = "Distribution of Listings by Borough",
        x = "Borough",
        y = "Number of Listings"
    ) +
    theme_minimal() +
    coord_flip() +
    theme(
        axis.text.y = element_text(size = 12),
        legend.position = "none"
    )
```

This bar chart shows the number of Airbnb listings for each NYC borough. Most of the listings are in Manhattan. Staten Island and Bronx had the lowest number of ratings, implying they are less popular for short-term rentals. 

```{r}
# Count listings by neighborhood and select top 10
top_neighborhoods <- airbnb %>%
    count(neighbourhood, sort = TRUE) %>%
    top_n(10, n) %>%
    arrange(desc(n))

# Calculate percentages for labeling
top_neighborhoods <- top_neighborhoods %>%
    mutate(percentage = round((n / sum(n)) * 100, 1))

# Custom color palette
nyc_palette <- c(
    "Bedford-Stuyvesant" = "#A73535", "Bushwick" = "#FFD700",
    "Chelsea" = "#6B6B6B", "Crown Heights" = "#1C4E80",
    "Harlem" = "#2E8B57", "Hell’s Kitchen" = "#FF5733",
    "Midtown" = "#3498DB", "Upper East Side" = "#8E44AD",
    "Upper West Side" = "#E67E22", "Williamsburg" = "#D35400"
)

# Top 10 neighborhoods
ggplot(top_neighborhoods, aes(x = reorder(neighbourhood, n), y = n, fill = neighbourhood)) +
    geom_bar(stat = "identity", width = 0.6, color = "black") +
    scale_fill_manual(values = nyc_palette) +
    labs(
        title = "Top 10 Most Popular Neighborhoods for Airbnb Listings",
        x = "Neighborhood",
        y = "Number of Listings"
    ) +
    theme_minimal() +
    coord_flip() +
    theme(
        axis.text.y = element_text(size = 10),
        legend.position = "none"
    )
```
This bar chart shows the top 10 neighborhoods in NYC with the highest number of listings, emphasizing the most popular areas. Note that Bedford-Stuynesant has the most listings with over 1500 listings. 